---
title: "Clean and Visualization of Palmer Penguins Dataset"
author: "By Yash Sharma"
date: ''
output: html_document
---

# Introduction
This is an exploratory analysis of [The Palmer Penguins Dataset from ASHWANI RATHEE on Kaggle](https://www.kaggle.com/datasets/ashkhagan/palmer-penguins-datasetalternative-iris-dataset).  

Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network.

# Palmer Penguins Dataset

## Setting up my environment
Notes: Setting up my environment by load libraries of `tidyverse` and `palmerpenguins` packages:

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Inspect & Clean Data
For Preview the dataset we use the `head()` function which returns the columns and first several rows of data:

```{r}
head(penguins)
```

In addition to `head()` function there are few several function for preview the dataset. for example `str()` function will provide summaries of each column in your data arranged horizontally:

```{r}
str(penguins)
```
In addition to `str()` function there are few several function for preview the dataset. for example `glimpse()` function will provide summaries of each columns:

```{r}
glimpse(penguins)
```

To find out what columns I have in our data frame. `colnames()` function help to preview the columns names in our dataset:

```{r}
colnames(penguins)
```

## Manipulating our Data
Filter our dataset which contains only `Adelie` species 

```{r}
filter(penguins, species == "Adelie")
```

Arrange our dataset in descending order

```{r}
arrange(penguins, -bill_length_mm)
```

Now we want to group species in our dataset for this we use `group_by()` function to group species:

```{r}
group_by(penguins,species)
```

If we want to filter,arrange and group_by our dataset with one line of code than we use pipe function that help our to filter,arrange and group_by our dataset at one time:

```{r}
penguins %>%
filter(island == "Torgersen") %>%
arrange(-bill_length_mm) %>%
group_by(species)
```

Now, we use `drop_na()` function to eliminate blank cells in our dataset:

```{r}
penguins %>%
filter(island == "Torgersen") %>%
arrange(-bill_length_mm) %>%
group_by(species) %>%
drop_na()
```

To create new variables in your data frame, we can use the `mutate()` function. This will make changes to the data frame, but not to the original data set you imported. That source data will remain unchanged:

```{r}
penguins %>%
filter(island == "Torgersen") %>%
arrange(-bill_length_mm) %>%
group_by(species) %>%
drop_na() %>%
mutate(flipper_length_m = flipper_length_mm / 1000)
```

Now, we use `select()` funtion to select fwe columns from our dataset:

```{r}
penguins %>%
select(species,island,bill_length_mm,flipper_length_mm) %>%
filter(island == "Torgersen") %>%
arrange(-bill_length_mm) %>%
group_by(species) %>%
drop_na()
```

Now, we use `summariize()` function for summarize our data eg: max,min,average,standard deviation of flipper length:

```{r}
penguins %>%
drop_na() %>%
summarize(maximum=max(flipper_length_mm),
          minimum=min(flipper_length_mm),
          average=mean(flipper_length_mm),
          standard_deviation=sd(flipper_length_mm))
```

## Visualizations
Here we will go through a series of visualizations (scatter plot,smooth line,jitter plot,line,violin,boxplot,bar,histogram) :-

### Flipper and Body Mass 
Here, we plot flipper length against body mass(except in bar and histogram chart):

```{r}
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_point(color="purple")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_smooth(method="loess",color="purple")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_jitter(color="purple")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_line(fill="black")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_violin(fill="black")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_boxplot()
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_bar(color="purple")
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_histogram(color="purple")
```

### Flipper and body mass by species
Here, we plot flipper length against body mass and look at the breakdown by species(except in bar and histogram chart):

```{r}
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_point(aes(shape=species))
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_smooth(method="loess")
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_jitter(aes(shape=species))
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_line(aes(shape=species))
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_violin(aes(shape=species,fill=species))
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_boxplot(aes(shape=species))
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_bar(aes(shape=species,color=species))
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_histogram(aes(shape=species))
```

### Flipper and body mass by species and sex
Here, we plot flipper length against body mass and look at the breakdown by species and sex(except in bar and histogram chart):

```{r}
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_point(aes(color=species,shape=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_smooth(method= "loess",aes(color=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_jitter(aes(color=species,shape=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_line(aes(color=species,shape=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_violin(aes(color=species,shape=species,fill=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm,y=body_mass_g)) + geom_boxplot(aes(color=species,shape=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_bar(aes(shape=species,color=species)) + facet_wrap(~sex)
ggplot(data=penguins,aes(x=flipper_length_mm)) + geom_histogram(aes(shape=species,color=species)) + facet_wrap(~sex)
```


### Add Title, Subtitle and Caption 
Now, we add a descriptive title, subtitle and Caption to our plot(except in bar and histogram chart):

```{r}
ggplot(data=penguins)+geom_point(mapping=aes(x=flipper_length_mm,y=body_mass_g,shape=species,color=species))+ labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_smooth(method="loess",mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species)) + labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_jitter(mapping=aes(x=flipper_length_mm,y=body_mass_g,shape=species,color=species))+ labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_line(mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species)) + labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_violin(mapping=aes(x=flipper_length_mm,y=body_mass_g,shape=species,fill=species)) + labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_boxplot(mapping=aes(x=flipper_length_mm,y=body_mass_g,color=species)) +labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_bar(mapping=aes(x=flipper_length_mm,color=species)) + labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
ggplot(data=penguins)+geom_histogram(mapping=aes(x=flipper_length_mm,color=species)) +labs(title="Palmer_Penguins:Flipper_Length_VS_Body_Mass", subtitle="Three Penguins Species", caption = "Created By Yash Sharma")
```

